---
title: "EDU DATA"
output: pdf_document
date: "2024-10-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyverse)
```

# Target Audience

target quantifies how many strategies and plans are aimed at each recipient
* each plan can have multiple strategies 
* each plan/strategy can target multiple recipients
* edu_target_audience is who the plan/strategy targets
* n_strategies is how many dinstinct action items target a recipient
* n_plans is how many MPA plans target a specified recipient across all action items
* percent_of_strategies is the percent of strategies that target a recipient
* percent_of_plans is the percent of plans that target a recipient
```{r}
target <- orig_list %>% 
  distinct(obs_id, plan_id, edu_target_audience) %>% 
  separate(edu_target_audience, into = paste0("part_", 1:8), sep = ";", fill = "right", extra = "merge") %>%
  mutate(across(everything(), str_trim)) %>% 
  pivot_longer(part_1:part_8, values_to = "edu_target_audience", names_to = NULL, values_drop_na = T) %>% 
  group_by(edu_target_audience) %>% 
  summarize(n_strategies = length(unique(obs_id)),
            n_plans = length(unique(plan_id))) %>% 
  mutate(percent_of_strategies = n_strategies/length(unique(orig_list$obs_id))*100,
         percent_of_plans = n_plans/length(unique(orig_list$plan_id))*100) 
```

## plots
```{r}
library(ggplot2)
target_barplot <- ggplot(target, aes(x = edu_target_audience)) +
  geom_bar(aes(y = n_strategies, fill = "Numbner of Strategies"), stat = "identity", position = "dodge") +
  geom_bar(aes(y = n_plans, fill = "Number of Plans"), stat = "identity", position = "dodge") +
  labs(title = "Counts of Stategies and Plans by Educational Target Audience",
       x = "Educational Target Audience",
       y = "Count",
       fill = "type") +
  theme(axis.text.x = element_text(angle = 45, size = 5, hjust = 1))
```

# EDU WHAT
* edu_what is social strategies and action steps of each overarching educational strategy
* plans can have multiple education strategy action steps
```{r}
edu_what <- orig_list |> 
  distinct(obs_id, plan_id, edu_what) |> 
  group_by(edu_what) |> 
  summarize(n_strategies = length(unique(obs_id)),
            n_plans = length(unique(plan_id))) |> 
  mutate(percent_of_strategies = n_strategies/length(unique(orig_list$obs_id))*100,
                percent_of_plans = n_plans/length(unique(orig_list$plan_id))*100) |> 
  rename(strategy_type = edu_what)
```

## plots
```{r}
ggplot(edu_what, aes(x = strategy_type)) +
  geom_bar(aes(y = n_strategies, fill = "Number of Strategies"), stat = "identity", position = "dodge") +
  geom_bar(aes(y = n_plans, fill = "Number of Plans"), stat = "identity", position = "dodge") +
  labs(title = "Counts of Strategies and Plans by Educational Strategy",
       x = "Educational Strategy",
       y = "Count",
       fill = "type") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 5), 
        plot.margin = margin(10, 10, 20, 30))
```

